Classless Java: Tuning Java Interfaces

### 2015-11-29 paper/sections/Translation.tex
Revise translation with @ObjWeak added.

\subsection{Translation Function}${}_{}$\\*

\noindent$\begin{array}{l}
\InTextDef{18em}{[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
}{
\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
\methods\ \methods' \cC
}\\
\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \methods'=\ofMethod(\C_0) \cup \otherMethod(\C_0,\methods)}
\end{array}$

To translate an annotated interface, we add the \Q@of@ method, and then we add some other methods.
However, first of all we check if the interface is valid for annotation:

@@

%To complete our formal definition we would need to add setters and \Q@with@.
Moreover, we check that the method \Q@of@ is not already defined by the user.
In our simplified formalization we consider this to be just an error.
In our prototype we keep overloading into account, and so we check that an \Q@of@ method with the same signature of the one we would generate is not already present.\marco{Do we check it?}

@@

\subsection{$\ofMethod$}${}_{}$\\*
We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory. To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.

@@

The function $\fieldsFunc(\C_0)$ (formally defined later) denotes all the fields in the current interface.
For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
  \begin{itemize}
   \item $\m_i$ is the field name, and have type $\C_i$.
   \item $\m_i$\QM{()} is the getter, that just return the current field value.
   \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
    the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This correspond to the expressions $\es_i$.
\item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modelling overloading.
%   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
   %\item To complete our generation, we need to generate setters, fluent setters and the with method.
   %\item \marco{should we just formalize setters?}
   \end{itemize}

\subsection{Other auxiliary functions}${}_{}$\\*

@@

As you can see above, \Q@with-@ and setter methods are generated if needed.
We can discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. Note that we do not need to check if its header is a subtype of what we would generate, this is ensured by $\valid(\C_0)$.

\noindent$\begin{array}{ll}
\InTextDef{20em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)}{
 \isWith(\mBody(\QM{with#}\m, \C_0))}\\
\InTextWith{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
\InTextDef{20em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0))}\\
\InTextWith{
\C\ \m\QM{();}\in \fieldsFunc(\C_0)
\mbox{ and } \QM_\m\notin\dom(\methods)}\\
%&\C_0\ \QM{clone}\oR\cR\QM;\in
%\otherMethod(\C_0)
%\\&  \mbox{ iff }
%\isClone(\mBody(\QM{clone}, \C_0))
%\mbox{ and } \QM{clone}\notin\dom(\methods)\\
\end{array}$

Other methods that we need to generate in the interface are \Q@with-@ and setters. %\Q@clone@.
%A complete formalization would also generate the \Q@with@.
This is needed only if we need to refine the return type.
To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
 is required by $\C_0$, but is not already present in the methods directly declared in $\C_0$.

\noindent$\begin{array}{ll}
\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
\isField(\method)\mand\
%}\\\InTextWith{
\method=\mBody(\_,\C_0)
}\\

\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
\mnot\ \specialName(\m)}\\
\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
 \mand\ \mnot\ \specialName(\m)}\\
\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
 \mand\ \mnot\ \specialName(\m)}\\

>>>>

\hl{Furthermore, we introduce \textbf{@ObjWeak}, which is only an annotation for illustrating the translation, hence is not implemented in Lombok. By this, the real translation is divided into two parts: \textbf{@Obj} automatically refines the return types only, and \textbf{@ObjWeak} generates the constructor method $\QM{of}$. In the source code, the only annotation \textbf{@Obj} combines both of the work and finishes the translation at a time.}

\subsection{Translation Function}

\noindent$\begin{array}{l}
\InTextDef{18em}{[\![\mixinAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
}{
\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
\methods\ \methods' \cC
}\\
\InTextWith{\methods'=\otherMethod(\C_0,\methods)}
\end{array}$

\hl{To translate an interface annotated by \mixin, we add some other methods for type refinement, then add a new annotation \textbf{@ObjWeak} to the interface.}

\noindent$\begin{array}{l}
\InTextDef{18em}{[\![\weakAnn\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC \methods\ \cC ]\!]
}{
\emptyset\ \QM{interface}\ \C_0\ \QM{extends}\ \Cs\ \oC
\methods\ \methods' \cC
}\\
\InTextWith{\valid(\C_0),\QM{of}\notin\dom(\methods) \mbox{ and } \methods'=\ofMethod(\C_0)}
\end{array}$

\hl{To translate an interface annotated by \textbf{@ObjWeak}, we add the $\QM{of}$ method.}
However, first of all we check if the interface is valid for annotation:

@@

%To complete our formal definition we would need to add setters and \Q@with@.
Moreover, we check that the method \Q@of@ is not already defined by the user.
In our simplified formalization we consider this to be just an error.
In our prototype we keep overloading into account, and so we check that an \Q@of@ method with the same signature of the one we would generate is not already present.\marco{Do we check it?}\haoyuan{Not sure if it is $\QM{of}\notin\dom(\methods)$ or $\QM{of}\notin\dom(\C_0)$. It is an error if a static method overrides an instance method.}

@@

\subsection{$\otherMethod$}
The definition of $\otherMethod(\C_0,\methods)$ is as follows:

\haoyuan{I think $\QM{with#}\m\notin\dom(\methods)$ is not enough. The withM method with return type $I_0$ can also be defined already in a super interface. Like:
\[\begin{array}{l}\QM{
interface A}\ \oC\QM{int x(); B withX(int x);}\cC\\
\mixinAnn\ \QM{interface B extends A}\ \oC\cC\end{array}\]}

\noindent$\begin{array}{ll}
\InTextDef{20em}{\C_0\ \QM{with#}\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)}{
 \isWith(\mBody(\QM{with#}\m, \C_0), \C_0)}\\
\InTextWith{\C\ \m\QM{();}\in \fieldsFunc(\C_0)
\mbox{ and } \QM{with#}\m\notin\dom(\methods)}\\
\InTextDef{20em}{\C_0\ \QM_\m\oR \C\ \QM{_val}\cR\QM;\in
\otherMethod(\C_0,\methods)}{\isSetter(\mBody(\QM_\m, \C_0), \C_0)}\\
\InTextWith{
\C\ \m\QM{();}\in \fieldsFunc(\C_0)
\mbox{ and } \QM_\m\notin\dom(\methods)}\\
\end{array}$

The methods that we generate in the interface are \Q@with-@ and setters. %\Q@clone@.
%A complete formalization would also generate the \Q@with@.
We discover if there is the need of generating such methods by checking if the method is unimplemented in $\C_0$. This is needed only if we need to refine the return type.
To discover if this is the case, we check if such \Q@with-@ or setter %\Q@clone@
 is required by $\C_0$, but is not already present in the methods \hl{directly declared in} $\C_0$.

\noindent$\begin{array}{ll}
\InTextDef{24ex}{\method\in\fieldsFunc(\C_0)}{
\isField(\method)\ \mand\
%}\\\InTextWith{
\method=\mBody(\_,\C_0)
}\\

\InTextDef{24ex}{\isField(\C\ \m\oR\cR\QM;)}{
\mnot\ \specialName(\m)}\\
\InTextDef{24ex}{\isWith(\C'\ \QM{with#}\m \oR \C\ \x\cR\QM;, \C_0)}{
\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
\ \mand\ \mnot\ \specialName(\m)}\\
\InTextDef{24ex}{\isSetter(\C'\ \QM_\m \oR \C\ \x\cR\QM;, \C_0)}{
\C_0 <: \C', \mBody(\m, \C_0) = \C\ \m\oR\cR\QM;
\ \mand\ \mnot\ \specialName(\m)}\\

\subsection{$\ofMethod$}
We now formally define $\ofMethod$, the function that generates the method \QM{of}, that behaves like a factory. To avoid boring digressions about well known ways to find unique names, for the sake of this formalization we assume that no-args methods do not start with underscore, and we prefix method names with underscore to obtain valid  parameter names.

@@

The function $\fieldsFunc(\C_0)$ (defined before) denotes all the fields in the current interface.
For methods inside the interface with the form $\C_i\ \m_i$\QM{();}
  \begin{itemize}
   \item $\m_i$ is the field name, and have type $\C_i$.
   \item $\m_i$\QM{()} is the getter, that just return the current field value.
   \item if a method \Q@with#@$\m_i$ is required, then it is implemented by calling the \Q@of@ method using
    the current value for all the fields except for $\m_i$. Such new value is provided as parameter. This correspond to the expressions $\es_i$.
\item \QM_$\m_i$\QM($\C_i\ $\QM{ _val)} is the setter. In our prototype we use name $\m_i$, here we use the underscore to avoid modelling overloading.
%   \item similarly, for the \Q@clone@ method, \Q@of@ is called using the current value for all the fields.
   %\item To complete our generation, we need to generate setters, fluent setters and the with method.
   %\item \marco{should we just formalize setters?}
   \end{itemize}

The auxiliary functions are defined below:\\*

@@

\subsection{Other features}
